---
output: html_document
---

# Fine-Tuning Publication-Quality ggplots

```{r setup-3, include = FALSE, purl = FALSE}
library(tidyverse)
raw_berry_data <- read_csv("data/clt-berry-data.csv")
raw_cider_data <- read_csv("data/CiderDryness_SensoryDATA.csv")
load("data/workshop-data.RData")
load("data/goal-plots.RData")
knitr::opts_chunk$set(fig.align = "center", warning = FALSE, message = FALSE)
```

The plots we've been making so far are fairly straightforward demonstrations. At the very end of Chapter 2, we briefly showed you the code and results for these two plots:

```{r pretty plot examples, fig.width = 7}
p1_berry_penalty
p2_ca_cider_cata
```

We're going to spend the rest of the workshop running through the tips and tricks we used to make these look polished. It's not an exhaustive list of the things you can change or tweak, though, so more than anything, we hope you walk away with the ability to identify what parts of a plot you want to change and the skills to figure out how to change it.

But first, something that everyone will need for almost every plot:

## Exporting and Saving Plots

It may seem like a weird time to be talking about saving plots, when we haven't gotten to the "fine-tuning" yet, but you've already seen a few examples of things that change when the aspect ratio of the plot or the monitor you're viewing it on changes slightly. It is basically impossible to export a `ggplot` at the resolution needed for publishing without changing the aspect ratio, relative sizes, or space between some elements from what you're looking at in the Plots pane or the `.Rmd` chunk output.

It's good practice to export your plot as an image (or pdf or knit-together document or however else you plan to export it) and re-export it periodically as you go through fine-tuning.

This is also, probably, the most important part of this chapter. You will have to get every plot that you publish out of `R` somehow, after all!

### Exporting Images with `ggsave()` {- #ggsave}
You can, as we've already discussed, save a `ggplot` object in a `.rds` file. But that won't let you put it into your powerpoint, or manuscript, or take it to the printer. You need an image file. The exact *type* of image will depend on the other software you're using for your report, presentation, paper, book, etc. 

The easiest way to reproducibly save plots, so that all of your export options are in your code and you might be able to recreate it on a different computer, is with the function `ggplot2::ggsave()`, which works *similarly* to the `write_*()` functions and `save()`. You give it a file name to save to, relative to the current working directory, and then the variable that has your ggplot.

```{r ggsave}
ggsave("img/penalty-lift.jpeg", p1_berry_penalty)
```

We saved our image as a JPEG file, because JPEG is one of the formats preferred by [Food Quality and Preference](https://www.sciencedirect.com/journal/food-quality-and-preference/publish/guide-for-authors#toc-51). `ggsave()` supports a wide variety of image formats and will automatically use the file type corresponding to the file extension (the stuff after the `.`) in the `filename` argument.

Note that `ggsave()` is like making a `.pdf` version of your working documents: you will not be able to read the plot images into `R` for editing anymore, no matter which format you save in.

We can also specify the `width` and `height` of our plot in inches, centimeters, or several other possible `units`. Aside from the **aspect ratio** of height to width, the size also affects how large various elements *look*, since `geom_*()` and text sizes are ultimately defined as fractions of inches.

Start by exporting your file at different `width` and `height` combinations until it looks legible, with readable text and other elements.

```{r ggsave-sized}
ggsave("img/penalty-lift-6x4.jpeg", p1_berry_penalty,
       width = 6,
       height = 4,
       units = "in") #inches
```

And then figure out the `dpi`, or number of dots per inch, needed to have a file that will print well. *Do not* just use the DPI you want to print at, especially if you're planning to use an image for a poster.

The FQAP Guide for Authors also tells us that it wants "combination figures", which will include most graphs, at or above 500 dpi. A double column figure is about 7.5 inches or 190mm wide, so we need the image to be $7.5 \times 500 = 3750$ pixels wide. Since our nice plot had `height = 6`, we'll set the `dpi` so that $height \times dpi \geq 3750$.

```{r fqap-resolution}
ggsave("img/penalty-lift-double-column-width.jpeg", p1_berry_penalty,
       width = 6,
       height = 4,
       units = "in", #inches
       dpi = 650) #slightly above 3750 / 6. Always err on the side of slightly too big!!
```

We go into a lot more detail on common file types and image resolution math in the [Appendix](#ggformat).

## Fine-tuning bar plots & facets with different categorical variables

We'll start with the penalty analysis plot.

`p1_berry_penalty` uses some tricks from the `stringr` and `tidytext` packages in order to give us easy-to-read labels. Using the existing column names and variable-codes in our original data to make a first draft of a plot, it would've looked more like this:

```{r worse penalty analysis example, fig.width = 7}
berry_penalty_analysis_data %>%
  ggplot(mapping = aes(x = cata_variable, y = penalty_lift)) +
  geom_col(aes(fill = penalty_lift), color = "white", show.legend = FALSE) + 
  facet_wrap(~berry,) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90))
```

Which we're showing because we have seen similar plots published in journal articles, with the difficult-to-read axes, underscores, redundant axis labels, and all. We can make this more readable by reordering the CATA attributes, shortening and reformatting the labels, and possibly by removing some extraneous elements like the `cata_variable` label. These are common steps that make a huge difference.

### Horizontal text with coord_flip()

We've used `coord_flip()` a few times already, and it's pretty straightforward to use--you just have to add it to a `ggplot()` call and it'll make your `x` axis the vertical axis and your `y` axis the horizontal.

```{r coord_flip}
berry_penalty_analysis_data %>%
  ggplot(mapping = aes(x = cata_variable, y = penalty_lift)) +
  geom_col(aes(fill = penalty_lift), color = "white", show.legend = FALSE) + 
  facet_wrap(~berry, nrow = 1) +
  theme_classic() +
  coord_flip()
```

Even though the CATA attributes are still cramped and this isn't the best use of space, they're much easier to read now. If you've only got one categorical axis, it's almost always going to be more readable to have it be the *vertical axis* so that the text is horizontal.

### Ordered Categorical Variables

Many of the figures we've made so far have had an axis with a categorical variable. Have you figured out how `ggplot2` orders the levels of categorical variables? If you have noticed, it's likely because it's in a different order than the one we'd like.

```{r default categorical axis in ggplot2}
berry_penalty_analysis_data %>%
  ggplot(mapping = aes(x = cata_variable, y = penalty_lift)) +
  geom_col(aes(fill = penalty_lift), color = "white", show.legend = FALSE) + 
  facet_wrap(~berry, nrow = 1) +
  coord_flip() + 
  theme_classic()
```

The CATA attributes are in alphabetical order (with the start of the alphabet the closest to 0). This is how `ggplot2` treats all `character` variables, and you can exert some control over the ordering by turning the variable into an ordered `factor`.

#### Specifying Ordinal Variables as Factors

You can order variables by hand, if there's a particular order you have in mind:

```{r manually making ordered factors}
berry_penalty_analysis_data %>%
  filter(str_detect(cata_variable, "taste")) %>%
  mutate(cata_variable = factor(cata_variable,
                                levels = c("taste_fruity",
                                           "taste_melon", "taste_peachy", "taste_grapey", "taste_grape",
                                           "taste_berry", "taste_cherry",
                                           "taste_citrus", "taste_lemon",
                                           "taste_tropical",
                                           "taste_candy", "taste_caramel",
                                           "taste_green", "taste_grassy", "taste_piney", "taste_minty",
                                           "taste_earthy", "taste_fermented",
                                           "taste_cinnamon", "taste_clove",
                                           "taste_floral",
                                           "taste_none"))) -> berry_penalty_manual_factors

berry_penalty_manual_factors %>%
  ggplot(mapping = aes(x = cata_variable, y = penalty_lift)) +
  geom_col(aes(fill = penalty_lift), color = "white", show.legend = FALSE) + 
  facet_wrap(~berry, nrow = 1) +
  coord_flip() + 
  theme_classic()
```

Note that the attribute you list *first* when you're specifying the `levels` will become 1, then 2, then 3. With `coord_flip()`, that puts it at the bottom of the plot.

```{r showing the order of the factor}
berry_penalty_manual_factors %>%
  distinct(cata_variable) %>%
  mutate(variable_number = as.numeric(cata_variable))
```

This gives us control, but it's pretty annoying to write out for large lists of attributes, and you have to be sure the spelling and capitalization match exactly. Often, like with the penalty analysis plots, what we actually want to do is order the Attributes in terms of some other numerical variable, like frequency or size of penalty.

One way is to `arrange()` the data the way you want it and then use that order to specify the levels.

```{r using another variable to order a factor}
berry_penalty_analysis_data %>%
  # Counting the number of times each attribute is used across all products:
  group_by(cata_variable) %>%
  mutate(variable_count = sum(count)) %>%
  ungroup() %>%
  # Arranging from least-to-most used:
  arrange(variable_count) %>%
  # Converting to a factor, so the least-used will be 1st, then the next:
  mutate(cata_variable = factor(cata_variable, levels = unique(cata_variable),
                            ordered = TRUE),
         variable_number = as.numeric(cata_variable)) -> berry_penalty_frequency_factors

#Now the plot:
berry_penalty_frequency_factors %>%
  ggplot(mapping = aes(x = cata_variable, y = penalty_lift)) +
  geom_col(aes(fill = penalty_lift), color = "white", show.legend = FALSE) + 
  facet_wrap(~berry, nrow = 1) +
  coord_flip() + 
  theme_classic()
```

#### Facets with Different Category-Orders

You'll notice that our reordered categorical axes still have the same order across all of the plots. This would be true even if we used the within-product sums already in the `count` column to calculate `level`s. The order is based on factor levels, which are fixed within each column: `Fresh_Apples` can't be "more than" `Dry` in one part of the `cata_variable` column and "less than" in another part.

On its own, `facet_wrap(..., scales = "free")` can drop unneeded attributes from plots, but it will still keep the same *order* of the attributes across all axes.

```{r free scales on ordered data}
berry_penalty_frequency_factors %>%
  ggplot(mapping = aes(x = cata_variable, y = penalty_lift)) +
  geom_col(aes(fill = penalty_lift), color = "white", show.legend = FALSE) + 
  facet_wrap(~berry, nrow = 1) +
  coord_flip() + 
  theme_classic()
```

If you have a faceted plot and you want each facet to have a different ordering of the terms, like in our big penalty analysis example, you'll have to use `tidytext::reorder_within()`, `tidytext::scale_*_reordered()`, *and* `facet_wrap(..., scales = "free")`, all at once:

```{r tidytext reordering within facets}
berry_penalty_analysis_data %>%
  mutate(cata_variable_clean = tidytext::reorder_within(cata_variable_clean,
                                                  by = count,
                                                  within = berry)) %>%
  ggplot(mapping = aes(x = cata_variable_clean, y = penalty_lift)) +
  geom_col(aes(fill = penalty_lift), color = "white", show.legend = FALSE) + 
  tidytext::scale_x_reordered() +
  facet_wrap(~berry, scales = "free", nrow = 1) +
  coord_flip() + 
  theme_classic()
```

### Making labels look okay: Powerful text manipulation with `stringr`

A good `R` variable or column name doesn't have any spaces or punctuation other than underscores (`_`) and dots (`.`), to avoid all those pesky backticks (`\``) in our code.

This is very different from what a good label in a plot looks like. You'll often want to make some sort of mass changes to column names or text variables before plotting, in order to address this.

The `stringr` package is a part of the `tidyverse`, so you have it already loaded whenever you run `library(tidyverse)`. It has a lot of useful functions for working with text (called "**str**ings" in many programming languages), mostly of the form `str_*()`. One thing you can do is change labels to uppercase, lowercase, "sentence case", or "title case" (first letter of each word capitalized), as appropriate:

```{r capitalizing with stringr}
berry_penalty_analysis_data %>%
  select(berry) %>%
  mutate(Upper = str_to_upper(berry),
         Title = str_to_title(berry)) # Capitalizes the first letter of each word
```

`str_replace()` and `str_replace_all()` are very useful for dealing with underscores or periods. You give it `string`, the text vector you want to modify (inside `mutate()`, a column name); then `pattern`, the character(s) you want to replace; then `replacement`, what you want to replace them with.

```{r replacing underscores with str_replace}
berry_penalty_analysis_data %>%
  mutate(cata_variable = str_replace_all(cata_variable, "_", ": "))
```

They can both be as long as you like. If `pattern = _` and there's more than one `_` in one of the `string`s, `str_replace()` will only replace the first one and `str_replace_all()` will replace them all.

```{r str_replace vs str_replace_all}
str_replace("long_text_with_many_underscores", "_", " ")
str_replace_all("long_text_with_many_underscores", "_", " ")
```

You can add multiple sets of `pattern`s and `replace`ments to `str_replace_all()` using a named list: `c("pattern1" = "replace1", "pattern2" = "replace2"`. This is useful to, for instance, fix multiple typos like "goodshapre". Unlike the `case_when()` example we showed before, you can fix typos that occur anywhere in the text, even if they're only parts of a word.

```{r fixing typos with str_replace}
berry_penalty_analysis_data %>%
  mutate(cata_variable = str_replace_all(cata_variable,
                                         c("shapre" = "shape",
                                           "appearane" = "appearance",
                                           "_" = " ")))
```

Be careful using short `pattern`s, because they will replace every example *even if it's only part of a word*.

```{r replacing parts of words}
#This can lead to unintentional side-effects
c("nocolor", "none", "cornonthecob", "anode") %>%
  str_replace_all("no", " NO ")

#Or it can be useful for fixing lots of similar problems all at once
berry_penalty_analysis_data %>%
  mutate(cata_variable = str_replace_all(cata_variable,
                                         c("not" = "not ",
                                           "good" = "good ",
                                           "uneven" = "uneven ",
                                           "_" = " ")))
```

So far, we've been replacing letters and underscores, which is what we have in our example data. You can also use `str_replace()` for periods (`.`), although you may be surprised when you first try:

```{r str_replace with periods}
str_replace_all("long.text.with.many.periods", ".", " ") # Replaces everything
str_replace_all("long.text.with.many.periods", "\\.", " ") # Replaces only dots
```

We have to **escape** the period (with an escaped backslash, technically, but for now just know that you must put two backslashes `\\` before special characters when using `stringr`). Because the `str_*` functions with a `pattern` can use **Regular Expressions** (or regex), the characters (`\`, `.`, `[`, `]`, `{`, `}`, `(`, `)`, `<`, `>`, `*`, `+`, `-`, `=`, `!`, `?`, `^`, `$`, and `|`) need to be **escaped** with two backslashes if you need to replace them.

Regex are extremely powerful tools for finding patterns in text, similar to the intuitive ways a human might recognize something like an email address, a measurement, or a parenthetical. We will not be talking about regex today, but if you want to see some examples and resources for learning how to use them, we've provided a short overview and links to some resources in the [Appendix](#regex).

It's also very common that you may have variable names in one column which actually represent multiple variables, and may want to, say, color-code the modality of the cata variables so that appearance attributes are one color and taste attributes are another.

There are a few versions of the `tidyr::separate_wider_*()` function that can, for example, split the `cata_variable` column into two columns, one with everything *before* the **delimiter** (here, an underscore) and one with everything *after* the delimiter. Be very careful using `separate_wider_delim()` like this if some rows may have more underscores (or whatever your delimiter is) than others.

```{r using separate to split one text column into multiple}
berry_penalty_by_modality <- berry_penalty_analysis_data %>%
  separate_wider_delim(cata_variable, "_",
                       names = c("mode", "variable"),
                       cols_remove = FALSE)

berry_penalty_by_modality

# Which would let you easily color-code your ggplot based on the new mode column

berry_penalty_by_modality %>%
  ggplot(aes(x = cata_variable, y = count, fill = mode)) +
  geom_col() +
  coord_flip()
```
(We'll get to the typo in just a second--you already know how to fix it, though!)

If you want to put multiple columns of text *together*, you can use either `tidyr::unite()` or `stringr::str_c()` (string combine).

```{r combining text columns}
berry_penalty_by_modality %>%
  mutate(cata_variable_2 = str_c(variable, mode, sep = " ")) %>%
  unite(variable, mode, col = "cata_variable_3", sep = " ")
```

You may have noticed that `berry_penalty_analysis_data` has both `cata_variable` and `cata_variable_clean`, and that the latter is shorter, which fits better on the page. You now know all of the tricks that we used to get from one to the other:

```{r using multiple string manipulations in practice}
berry_penalty_analysis_data %>%
  # Fixing typos
  mutate(cata_variable = str_replace_all(cata_variable, c("appearane" = "appearance",
                                                      "shapre" = "shape"))) %>%
  # Separating the modality from the attribute name
  separate_wider_delim(cata_variable, "_",
                       names = c("mode", "variable"),
                       cols_remove = FALSE) %>%
  # Shorthand for the taste and appearance variables
  mutate(mode = case_when(mode == "taste" ~ "(T)",
                          mode == "appearance" ~ "(A)")) %>%
  # Putting them back together
  unite(variable, mode, col = "cata_variable_clean", sep = " ")
```

### Removing Plot Elements

Plots are easiest to read when they have only the information that's actually relevant, so some labels and other elements that `ggplot2` adds by default can end up being extraneous and distracting.

We covered using `theme()` to remove elements with `element_blank()` (e.g., `+ theme(panel.grid = element_blank())` to get rid of the grid lines), but there are some exceptions to this rule.

When we're using `tidytext::reorder_within()` to order each individual facet from highest penalty to highest lift, any color-coding for the lift is redundant info and we don't really need a legend to spell it out:

```{r redundant legend}
berry_penalty_analysis_data %>%
  mutate(cata_variable_clean = tidytext::reorder_within(x = cata_variable_clean,
                                                        by = penalty_lift,
                                                        within = berry)) %>%
  ggplot(mapping = aes(x = cata_variable_clean, y = penalty_lift)) +
  geom_col(aes(fill = penalty_lift), color = "white") + 
  facet_wrap(~berry, scales = "free", nrow = 1) + 
  coord_flip() + 
  tidytext::scale_x_reordered()
```

We also don't really need the axis labels, since neither has units and anyone unfamiliar with a penalty-lift plot will need more explanation than an axis label can really provide, but `cata_variable_clean` is especially redundant.

```{r removing the legend}
berry_penalty_analysis_data %>%
  mutate(cata_variable_clean = tidytext::reorder_within(x = cata_variable_clean,
                                                        by = penalty_lift,
                                                        within = berry)) %>%
  ggplot(mapping = aes(x = cata_variable_clean, y = penalty_lift)) +
  geom_col(aes(fill = penalty_lift), color = "white") + 
  facet_wrap(~berry, scales = "free", nrow = 1) + 
  coord_flip() + 
  tidytext::scale_x_reordered() +
  theme(axis.title.y = element_blank(),
        legend.position = "none")
```

We can remove these both with one call to `theme()`, but the axis title is removed with `axis.title.y = element_blank()` and the legend is removed with `legend.position = "none"`. And, yes, that *is* `axis.title.y`, referring to the same `cata_variable_clean` axis that `scale_x_reordered()` is calling the x-axis.

Some `ggplot2` functions will always refer to whatever you specified in `aes(x = ...)` as the x-axis, while some will refer to the *horizontal* axis as the x-axis, meaning that `coord_flip()` can make things confusing. It's not worth trying to memorize or find patterns--this is something you'll have to tweak by trial and error.

Sometimes, there are multiple ways of removing a plot element, and whichever one you remember is fine to use. One notable example that may save you some typing is the `labs()` function, which lets you label the `x` and `y` axes, the whole plot `title`, and so on with a slightly easier-to-remember interface than `theme()`. Setting any of the `labs()` arguments `= NULL` will remove it from the plot.

```{r removing the legend}
berry_penalty_analysis_data %>%
  mutate(cata_variable_clean = tidytext::reorder_within(x = cata_variable_clean,
                                                        by = penalty_lift,
                                                        within = berry)) %>%
  ggplot(mapping = aes(x = cata_variable_clean, y = penalty_lift)) +
  geom_col(aes(fill = penalty_lift), color = "white") + 
  facet_wrap(~berry, scales = "free", nrow = 1) + 
  coord_flip() + 
  tidytext::scale_x_reordered() +
  labs(x = NULL, y = "Penalty",
       title = "Penalty / Lift Analysis",
       subtitle = "displays the mean difference (within berries) for when a CATA variable is checked")
```

Note that, unlike `theme()`, `labs(x = ...)` refers to the same `x` as `aes(x = ...)`.

### Putting it all together

Now, at long last, we're ready to walk line-by-line through the example penalty analysis figure that we've just been copy-pasting so far in the workshop.

```{r final walkthrough of penalty analysis, fig.width = 7}
berry_penalty_analysis_data %>%
  mutate(cata_variable_clean = tidytext::reorder_within(x = cata_variable_clean,
                                                        by = penalty_lift,
                                                        within = berry)) %>%
  ggplot(mapping = aes(x = cata_variable_clean, y = penalty_lift)) +
  geom_col(aes(fill = penalty_lift), color = "white") + 
  facet_wrap(~berry, scales = "free", nrow = 1) + 
  tidytext::scale_x_reordered() + 
  coord_flip() + 
  theme_classic() + 
  scale_fill_gradient(low = "tan", high = "darkgreen") + 
  labs(title = "Penalty / Lift Analysis",
       subtitle = "displays the mean difference (within berries) for when a CATA variable is checked") +
  theme(axis.title = element_blank(),
        legend.position = "none")
```

## Fine-tuning biplots with different types of variables

`p2_ca_cider_cata` has a lot of different aesthetics, some of . Using the existing column names and variable-codes in our original data to make a first draft of a plot, it would've looked more like this:

```{r worse penalty analysis example, fig.width = 7}
ca_cider_coords %>%
  ggplot(mapping = aes(x = `Dim 1`, y = `Dim 2`, color = type)) + 
  geom_point() +
  ggrepel::geom_text_repel(mapping = aes(label = name)) +
  theme_bw() + 
  scale_color_manual(values = c("darkorange", "darkgreen"))
```

### Tweaking legends

If you want to remove *specific geoms* from the legend, then you have to adjust your `geom_*()` calls.

```{r excluding geoms from the legend}
ca_cider_coords %>%
  filter(type == "col") %>%
  ggplot(aes(x = `Dim 1`, y = `Dim 2`,
             label = name, color = modality)) -> ca_cider_colored

ca_cider_colored +
  geom_point() +
  ggrepel::geom_text_repel()

ca_cider_colored +
  geom_point() +
  ggrepel::geom_text_repel(show.legend = FALSE)
```

### Axis and plot labels based on variables

The `FactoMineR` plot we briefly showed in [Chapter 1](#pubquality) included the percentage of inertia explained by each dimension *in the axis labels*, which is extremely nice, even though it's difficult to adjust the `FactoMineR` plot directly.

The `FactoMineR::CA()` function gives us all of the information we need to add this to our ggplot, and in our data set it's saved in the `ca_cider$eig` matrix.

```{r matrix of eigenvalues and intertia from FactoMineR}
ca_cider$eig #the second column has the percentage of variance explained by each dimension
```

if you give `str_c()` multiple vectors with the same length, it'll combine the first element of the first vector with the first element of the second vector, and so on. So we could easily make a list of plot labels for each axis we have:

```{r list of axis names with str_c}
str_c(rownames(ca_cider$eig),
      " (", # you'll need to add the symbols yourself, including spaces
      round(ca_cider$eig[, 2], 1),
      "%)")

# Or you can do it one at a time:
str_c("Dimension 1 (",
      round(ca_cider$eig[1, 2], 1),
      "% of intertia)")
str_c("Dimension 2 (",
      round(ca_cider$eig[2, 2], 1),
      "% of intertia)")
```

To add it to the plot, we can use the `labs()` function, which lets you label the `x` and `y` axes, the whole plot `title`, and so on.

```{r making wordy axis labels}
ca_cider_coords %>%
  ggplot(mapping = aes(x = `Dim 1`, y = `Dim 2`, color = type)) + 
  geom_point() +
  ggrepel::geom_text_repel(mapping = aes(label = name)) +
  theme_bw() + 
  scale_color_manual(values = c("darkorange", "darkgreen")) +
  labs(x = str_c("Dimension 1, ", round(ca_cider$eig[1, 2], 1), "% of inertia"),
       y = str_c("Dimension 2, ", round(ca_cider$eig[2, 2], 1), "% of inertia"),
       subtitle = "Correspondence Analysis biplot (symmetric)",
       #if you want to insert a line break, you can use `\n`:
       title = "Effect of cider serving temperature on consumer sensory perception\nbased on CATA data")
```

### Setting up the Cartesian plane

You may have noticed that `ggplot2` doesn't give any particular weight to the origin, or the gridlines that pass through it, even though (0,0) is often meaningful and we may want to see which points are positive or negative at a glance. Removing all gridlines or using `theme_void()` removes the x- and y-axes as well!

The easiest way to address this is just to add two lines that are thicker/darker/different at $y = 0$ and $x = 0$, which we can do by adding more `geom_*()`s...

```{r using geom_vline and geom_hline at the end}
ca_cider_coords %>%
  ggplot(mapping = aes(x = `Dim 1`, y = `Dim 2`, color = type)) + 
  geom_point() +
  ggrepel::geom_text_repel(mapping = aes(label = name)) +
  theme_void() +
  geom_vline(xintercept = 0, color = "grey", linewidth = 1) +
  geom_hline(yintercept = 0, color = "grey", linewidth = 1)
```

But if you aren't careful and add them *after* your other geoms, it will add them *on top* of the rest of the plot. In general, you should go ahead and add your axis lines right after calling `ggplot()`, before anything else, so that they're drawn on first and every other `geom_*()` goes on top of it.

```{r using geom_vline and geom_hline at the beginning}
ca_cider_coords %>%
  ggplot(mapping = aes(x = `Dim 1`, y = `Dim 2`, color = type)) +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = 0) + 
  geom_point() +
  ggrepel::geom_text_repel(mapping = aes(label = name)) +
  theme_void()
```

You can also make your plot more intuitive to read by ensuring that 1 unit of the x-axis is the same distance as one unit on the y-axis, using `coord_equal()`.

```{r using geom_vline and geom_hline at the beginning}
ca_cider_coords %>%
  ggplot(mapping = aes(x = `Dim 1`, y = `Dim 2`, color = type)) +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = 0) + 
  geom_point() +
  ggrepel::geom_text_repel(mapping = aes(label = name)) +
  theme_bw() +
  coord_equal()
```

This will often make plots resulting from SVD wider than they are tall, since the 1st dimension will definitionally explain more variation in the data.

### Using only part of the dataset for each geom

